Etox-Base: A tool to derive exposure endpoints from chemical test results

### 2019-02-08 article/article.tex
Updates from Overleaf

\citep{dowle_data.table_2018}.

\subsection*{Future of the project}

@@

The code for processing the data is stored in a Github reopsitory (\git{}). All processing scripts were written in R version 3.4.4. PostgreSQL 9.5 has been used to build the data base. Hence, R together with its packages and PostgreSQL suffice to rebuild the process. The \app{} itself is stored in a second Github repostitory (\gitapp{}).
Since new versions of the \epa{} are published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are scheduled to be released regularly.

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
         Category       & Package       & Version   & Reference \\
         processing     & RCurl         & \\
         & 
    \end{tabular}
    \caption{List of R packages used in compiling the data}
    \label{tab:rpackages}
\end{table}

@@

The \etoxbase{} can be accessed via an application here: \app{}. The user can filter among different test parameters \ref{tab:inputs}.

The \etoxbase{} is designed to support ecotoxicologists in assessing the risk of chemicals on the environment. With an increase of ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats, meaning accessible for humans as well as machines. Currently chemical risk assessment relies on test results of a few, well tested organism groups, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} reported in publications on test procedures <CITE>, compiled published data sets \citep{malaj_organic_2014} or data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}, whereby each source having its limitations: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to specific chemical groups, such as pesticides, biozides etc. The \etoxbase{}, making use of the \epa{} data aims to provide holistic aggregates (\ecfifty{}, NOEC, LOEC) of ecotoxicological tests results in adequate formats in order to support researchers to in the compilation of SSDs, TUs and other risk indicators.

@@

\bibliographystyle{apalike}
\bibliography{refs/references-etox-base,
              refs/references-etox-base-rpackages,
              refs/references-etox-base-rshinypackages}

>>>>

@@

The code for processing the data is stored in a Github reopsitory (\git{}). All processing scripts were written in R version 3.4.4. PostgreSQL 9.5 has been used to build the data base. Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process. The \app{} itself is stored in a second Github repostitory (\gitapp{}).
Since new versions of the \epa{} are published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are scheduled to be released regularly.

\begin{table}[]
    \centering
    \begin{tabular}{|l|l|l|l|}
        \hline
        Category        & Package       & Version   & Reference                     \\ \hline
        processing      & RCurl         & 1.95-4.11 & \citep{lang_rcurl_2018}       \\ \hline
        processing      & data.table    & 1.12.0    & \citep{dowle_data.table_2018} \\ \hline
        application     & shiny         & 1.1.0     & \citep{chang_shiny_2018}      \\ \hline
    \end{tabular}
    \caption{List of R packages used in compiling the data.\newline{}
    TODO: Automate version input and name generation - whole table\newline{}
    TODO2: debug chang\_shiny\_2018 reference}
    \label{tab:rpackages}
\end{table}

@@

The \etoxbase{} can be accessed via an application here: \app{}. The user can filter among different test parameters \ref{tab:inputs}. The \etoxbase{} is designed to support ecotoxicologists in assessing the risk of chemicals on the environment. With an increase of ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats, meaning accessible for humans as well as machines. Currently chemical risk assessment relies on test results of a few, well tested organism groups, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} reported in publications on test procedures <CITE>, compiled and published data sets \citep{malaj_organic_2014} or in data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}. Each data source having its limitations: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to specific chemical groups, such as pesticides, biozides etc. The \etoxbase{}, making use of the \epa{}, the largest publicly available collection of ecotoxicological test results, aims to provide holistic aggregates (\ecfifty{}, NOEC, LOEC) of ecotoxicological tests results in adequate formats in order to support researchers in fetching ecotoxicological data for the calculation of SSDs, TUs and other risk indicators. SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels \citep{posthuma_species_2002}. As an example

\begin{figure}
    \dentering
    \includegraphics[width=1\linewidth]{figures/ssd2}

\end{figure}

@@

\bibliographystyle{apalike}
\bibliography{refs/references-etox-base.bib,refs/references-etox-base-rpackages.bib,refs/references-etox-base-rshinypackages.bib}
% multiple refs have to be written w/o space
